Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-06-02 SecDev21/paper.tex
cutting text

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects. \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A frequent usage in postconditions is

@@

This states that the relevant fields of \spark{Sock} selected by \spark{Model} are not changed by the function call, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.

@@

The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, also has to be taken into account. The reason for this is two-fold. First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions. Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit. The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.

>>>>

The socket structure is shared between different functions and contains a large number of fields. Most of the fields are used in library parts that are not considered by this work, and thus, are not relevant for our verification. Thus we use \emph{ghost code} in SPARK to focus the verification effort on the relevant subset of the fields, which also makes the specifications easier to write and understand. This is a solution to the classical problem of specifying the \emph{frame condition} for a function, that is, the set of possibly modified objects. \emph{Ghost code} is used by GNATprove for analysis but is not being compiled in the final executable. More specifically, we define a \emph{model} of the actual socket as a ghost function that can be used in contracts: \spark{Model(Sock)} extracts the fields of interest from socket \spark{Sock}. A common usage is within a function postcondition:

@@

This states that the fields of \spark{Sock} used by \spark{Model} are not changed when calling the function, except for the field \spark{State} that is equal to \spark{TCP_STATE_CLOSED} after the call.

@@

The original \emph{CycloneTCP} library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, has to be taken into account. The reason for this is two-fold. First, although user-functions can trigger state transitions, the actual transitions are done by library functions that do not belong to the user-task set of functions. Second, other parts of the library can trigger state transitions during and between the user-task function calls, which can affect the intermediate and final states that a user-related function can exhibit. The TCP user API-related functions were fully translated to SPARK. SPARK bindings were also provided to most of the rest of the library's C functions to allow their invocation from the SPARK code.
